Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-15 paper/intro.tex
Minor updates to intro

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
%\footnote{The
%  views expressed are those of the authors and do not reflect the
%  official policy or position of the Defense Advanced
%  Research Projects Agency (DARPA) or the U.S. Government.}
We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.  
A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system. 
Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.  The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.  Cyber-resilient code implementing the verified model is then automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.

@@

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR-generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

>>>>

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
%\footnote{The
%  views expressed are those of the authors and do not reflect the
%  official policy or position of the Defense Advanced
%  Research Projects Agency (DARPA) or the U.S. Government.}
We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models for potential cyber vulnerabilities and suggest cyber-security requirements to mitigate them.
A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system. 
Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
Formal verification that the transformed system model satisfies its cyber requirements is accomplished via the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output. 
AGREE is a {\em compositional assume-guarantee} style model checker for AADL models that attempts to prove properties about one layer of the architecture using
properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.

@@

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect (and alert) unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the incorporation of these high-assurance components, the hardened system meets its cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming that the execution schedule of the deployed cyber-hardened system is as intended by the AADL model, and that the HAMR-generated communication fabric delivers messages between components as expected, the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
